---
title: "ECON 370 Final Project"
author: 'SD:2, 5, 17, 19, 28' 
date: "2022-11-8"
output:
  word_document: default
  pdf_document: default
---
# ellis
# ============================================
# Set-up
# ============================================
```{r}
library(readr)
library(mosaic)
library(data.table)
library("xlsx")
library(openxlsx)
library(mvtnorm)
library(tidyverse)
```

# ============================================
# Question 1 - Download Data
# ============================================

Note: r eval = FALSE prevents code chunk from running when data is already downloaded (the "switch" in our code).
Delete eval = FALSE for first run of code to download files.
```{r eval=FALSE} 
for (i in 1998:2010){
  
url1 <- paste("http://www.nber.org/hcris/265-94/rnl_rpt265_94_", i , ".csv",sep = '')
url2 <- paste("http://www.nber.org/hcris/265-94/rnl_nmrc265_94_", i, "_long.csv", sep = '')
url3 <- paste("http://www.nber.org/hcris/265-94/rnl_alpha265_94_", i, "_long.csv", sep = '')

destfile <- paste("./hcris_raw/rnl_rpt265_94_", i , ".csv",sep = '')
download.file(url1,destfile)
destfile <- paste("./hcris_raw/rnl_nmrc265_94_", i, "_long.csv", sep = '')
download.file(url2,destfile)
destfile <- paste("./hcris_raw/rnl_alpha265_94_", i, "_long.csv", sep = '')
download.file(url2,destfile)

}
```

# ============================================
# Question 2 - Cleaning The Data
# ============================================

## Read in raw data
```{r}
# Read in all CSVs
for (i in 1998:2010){
  
rpt_name = paste("rpt", i, sep = '')
nmrc_name = paste("nmrc", i, sep = '')
alpha_name = paste("alpha", i, sep = '')

rpt <- read_csv(paste("hcris_raw/rnl_rpt265_94_", i , ".csv",sep = ''), show_col_types = F)
nmrc <- read_csv(paste("./hcris_raw/rnl_nmrc265_94_", i, "_long.csv", sep = ''), show_col_types = F)
alpha  <- read_csv(paste("./hcris_raw/rnl_alpha265_94_", i, "_long.csv", sep = ''), show_col_types = F)

rpt$year = i
nmrc$year = i
alpha$year = i

nmrc$key = paste(nmrc$wksht_cd, nmrc$line_num, nmrc$clmn_num, sep = '')
alpha$key = paste(alpha$wksht_cd, alpha$line_num, alpha$clmn_num, sep = '')

assign(rpt_name, rpt)
assign(nmrc_name, nmrc)
assign(alpha_name, alpha)

}

# clean up environment
rm(rpt, nmrc, alpha, rpt_name, nmrc_name, alpha_name)

## DELETE THIS WHEN DONE: loading Alex's data for part 2
## note to all: use temp for cleaning data portion 
temp <- read_csv("hcris_panel_raw.csv")

variable_codes <- read.csv("variable_codes.csv")
#generate key for variable codes to parse
variable_codes$key = paste(variable_codes$worksheet, variable_codes$line, variable_codes$column, sep = '')

# fix space in data
variable_codes$key[35] = "S000001102300"
```

## Reformatting The Raw Data (current draft, merge isn't working. variable codes not lining up)
```{r eval=FALSE}
# create keys for alpha and nmrc data

# merge datasets
## vector memory exhausted error? 
## Question for alex: why are the datasets (nmrc and alpha) identical? 

# TO SOPHIA: I'm trying to parse through the dataframe to check for a variable_code match (to drop observations without a match)
toMatch = paste(variable_codes$key, collapse = "|")
result = filter(nmrc1999, grepl(toMatch, key))

# SOPHIA: problem is - there doesn't seem to be any matchs for the keys in any of the datafranes, 
sum(grepl(toMatch, nmrc1998$key))
sum(grepl(toMatch, nmrc1999$key))
sum(grepl(toMatch, nmrc2000$key))
sum(grepl(toMatch, nmrc2001$key))
sum(grepl(toMatch, nmrc2002$key))
sum(grepl(toMatch, nmrc2003$key))
sum(grepl(toMatch, nmrc2004$key))
sum(grepl(toMatch, nmrc2005$key))
sum(grepl(toMatch, nmrc2006$key))
sum(grepl(toMatch, nmrc2007$key))
sum(grepl(toMatch, nmrc2008$key))
sum(grepl(toMatch, nmrc2009$key))
sum(grepl(toMatch, nmrc2010$key))

sum(grepl(toMatch, alpha1998$key))
sum(grepl(toMatch, alpha1999$key))
sum(grepl(toMatch, alpha2000$key))
sum(grepl(toMatch, alpha2001$key))
sum(grepl(toMatch, alpha2002$key))
sum(grepl(toMatch, alpha2003$key))
sum(grepl(toMatch, alpha2004$key))
sum(grepl(toMatch, alpha2005$key))
sum(grepl(toMatch, alpha2006$key))
sum(grepl(toMatch, alpha2007$key))
sum(grepl(toMatch, alpha2008$key))
sum(grepl(toMatch, alpha2009$key))
sum(grepl(toMatch, alpha2010$key))


merge(x = nmrc1998, y = alpha1998, by = "key")
merge(x = nmrc1999, y = alpha1999, by = "key")
merge(x = nmrc2000, y = alpha2000, by = "key")
merge(x = nmrc2001, y = alpha2001, by = "key")
merge(x = nmrc2002, y = alpha2002, by = "key")
merge(x = nmrc2003, y = alpha2003, by = "key")
merge(x = nmrc2004, y = alpha2004, by = "key")
merge(x = nmrc2005, y = alpha2005, by = "key")
merge(x = nmrc2006, y = alpha2006, by = "key")
merge(x = nmrc2007, y = alpha2007, by = "key")
merge(x = nmrc2008, y = alpha2008, by = "key")
merge(x = nmrc2009, y = alpha2009, by = "key")
merge(x = nmrc2010, y = alpha2010, by = "key")

# helper function to check str length
check_count = function(){
count = 0
for (i in 1:dim(alpha1999)[1]){
  if (str_length(alpha1999$line_num[i]) != 5){
  count = count + 1
  print(i)
  }
}
print(count)
}

# #stackoverflow solution to vector limit problem: 
# library(usethis) 
# usethis::edit_r_environ()
```


## Cleaning The Data
```{r}
# 1

# 2

# 3

# 4

# 5

# 6

# 7

# 8

# 9

# 10

# 11
## what to do about NA values? for now: just removing them here 
sort(unique(temp$chain_identity))

# remove observations with chain_indicator == NA | chain_identity == NA
temp2 = subset(temp, chain_indicator == "Y" | chain_indicator == "N")
temp3 = subset(temp2,chain_identity != "NA")

temp3$chain_id = 0 # default for chain_indicator == "N"

# initialize chain_indicator variables 
for(i in 1:nrow(temp3)){
  if (temp3$chain_indicator[i] == "Y" & temp3$chain_identity[i] != "DAVITA" & temp3$chain_identity[i] != "FRESENIUS" ){
    temp3$chain_id[i] = 1
  }
  else if (temp3$chain_identity[i] == "DAVITA"){
    temp3$chain_id[i] = 2
  }
  else if (temp3$chain_identity[i] == "FRESENIUS"){
    temp3$chain_id[i] = 3
  }
}

## this line of code is not running, chain_identity not found 
temp3[grepl("DAVITA",chain_identity,ignore.case = TRUE),chain_identity := "DAVITA"]

# 12

```


# ============================================
# Question 3 - Analysis 
# ============================================
```{r}

```

